---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r loading_and_preprocessing}
dataFile <- read.table("activity.csv",sep=",",header=TRUE)
nona_dataFile <- na.omit(dataFile)
```

## What is mean total number of steps taken per day?

```{r mean}
library(plyr)
steps_per_day<-ddply(nona_dataFile,"date",numcolwise(sum))
hist(steps_per_day$steps)
mean_steps<-mean(steps_per_day$steps)
median_steps<-median(steps_per_day$steps)
```

The mean of steps per day is `r mean_steps`, and the median `r median_steps`

## What is the average daily activity pattern?

```{r activity}
steps_per_day$date <- as.Date( steps_per_day$date, '%Y-%m-%d')
require(ggplot2)
ggplot(steps_per_day, aes(date, steps)) + geom_line()
max_steps_interval<-steps_per_day$interval[steps_per_day$steps==(max(steps_per_day$steps))]
```

On average across all the days in the dataset, the 5-minute interval which contains the maximum number of steps is `r max_steps_interval`

## Imputing missing values

```{r na_values}
incomplete_cases<-nrow(dataFile)-nrow(na.omit(dataFile))

#Filling in all of the missing values in the dataset with 0

newDataFile<-dataFile 
newDataFile$steps[is.na(dataFile$steps)]<-0

#calculation with all values

steps_per_day_new<-ddply(newDataFile,"date",numcolwise(sum))
hist(steps_per_day_new$steps)
mean_steps_new<-mean(steps_per_day_new$steps)
median_steps_new<-median(steps_per_day_new$steps)
```

Total number of rows with NAs: `r incomplete_cases`.
The new mean of steps per day is `r mean_steps_new`, and the median `r median_steps_new`

## Are there differences in activity patterns between weekdays and weekends?

```{r differences}
steps_per_day_new$date <- as.Date( steps_per_day_new$date, '%Y-%m-%d')

#Create new column "day"

steps_per_day_new$day<-""

for (i in 1:nrow(steps_per_day_new)) { 
  wd<-weekdays(steps_per_day_new$date[i]) 
  if (wd=="sábado" || wd=="domingo") { 
    steps_per_day_new$day[i]<-"weekend" 
  } else { 
    steps_per_day_new$day[i]<-"weekday" 
  } 
}

qplot(date,steps,data=steps_per_day_new,facets=.~day)+ geom_line()
```
